6.867 Problem Set 1

### 2015-12-10 Project/Latex/Paper.tex
edits

\Authors{Dave Holtz\textsuperscript{1}, Jeremy Yang\textsuperscript{2}, Michael Zhao\textsuperscript{3*}} % Authors
\affiliation{\textsuperscript{1}\textit{dholtz@mit.edu}}
\affiliation{\textsuperscript{2}\textit{zheny@mit.edu}} 
\affiliation{\textsuperscript{3}\textit{mfzhao@mit.edu}}
\affiliation{\textsuperscript{*}\textit{Michael is also using this project to fulfill Final Project requirements for both 6.864 as well as 6.867. Dave and Jeremy are only using this project for 6.867.}}

@@

\Abstract{Why do some webpages receive massive numbers of page views? To determine how content drives viewership, we construct a unique dataset of all articles published by the New York Times (NYT) in August 2013. Our dataset is built from 2 major components, the NYT's internal web traffic data and article content data parsed from the NYT website. We use the internal web traffic data to accurately track the number of page views of each article as well as construct a set of robust control variables such as the desk and section of each article. To build content features, we use various machine learning and statistical natural language processing techniques on our parsed article content data and construct features such as article perplexity, sentiment, reading difficulty, and indicators that denote the presence of pictures, videos, etc. Additionally, we have access to the NYT's internal website traffic data. We feed all of our constructed features to into a predictive regression model. We find that our textual features modestly improve our predictive power. However, there are still many textual features we've not yet explored due to time constraints. We believe that adding these additional features may offer more significant improvements in performance.}

@@

In today's digital economy, many companies are very interested in attracting users to visit their websites in order to earn ad revenue. While many factors might motivate a user to visit a particular page, certainly one important factor is the content in that webpage. This paper explores the relationship between the content of a webpage and the number of page views it ultimately ends up receiving by constructing a unique dataset of all articles published by the New York Times (NYT) during August 2013. This dataset is built from two major components: the NYT's internal web traffic data and parsed NYT article content data.

Typically, a study such as ours tends to be very difficult to conduct as either accurate measures of viewership are unavailable\footnote{While oftentimes precise viewership data tends to be not available openly, oftentimes researchers use related observables, such as Facebook likes} or the feature extraction of the content is too challenging (for example Youtube), or or both. Fortunately, our access to the the NYT's internal web traffic data allows us to exactly measure the number of page views an article receives. The web traffic data is rather rich and also includes internal meta-data that we use to build various control features. Moreover, since we are working with mostly textual data, we are able to take advantage of recent advancements in machine learning and statistical NLP to do feature extraction on article text.

A similar study by Berger and Milkman (2012) \cite{berger2012makes} examines the relationship between content and word-of-mouth virality. They find that the emotional content of a NYT article is predictive of its virality. Using simple measures of an article's sentiment and emotionality, Berger and Milkman show that positive articles are more likely to show up on the New York Times "Most-Emailed" list. They also show that articles that evoke high physiological positive or negative arousal (such as awe or anger) tend to be more viral than articles that evoke deactivating emotions (sadness). We build on this study in two ways: first, we relate an article's content back to the number of page views it receives rather than its virality\footnote{Which companies arguably care more about since word-of-mouth virality is usually a means to increase page views}. Second, we employ more sophisticated machine learning feature extraction techniques to see if they work any better over their simple measures.
%------------------------------------------------

\section{Data}
\subsection{NYT Internal Web Traffic Data}
Our NYT internal web traffic dataset is a record of all individual user activity on the NYT website covering the period of April 3rd, 2013 to October 31st, 2013. This activity data is stored as individual lines of json and includes who (if available) accessed what page at what time. Overall, it is over 20 terabytes in size and contains over 3 billion page views. It's important to note that not all page views are content views, for example, some events that are also tracked are searches, or user account settings. Since the scope of this dataset is so large, we initially restrict this project to a single month, August 2013, with plans of including the entire dataset for future work.

We further constrain our dataset to only include articles or blogposts published during the month of August since these are the pieces of content that have machine parsable text data. For the sake of brevity, we use articles to refer to both articles and blogposts unless we explicitly state otherwise. We make a basic first pass through the data to simply to obtain a list of urls. After cleaning up the url data to make sure we had each url mapped to a unique piece of content, we were left with a total of 6682 urls. We then parse our dataset and aggregate the number of counts each url receives. In order to make the comparison between articles fairer since an article that's been out longer will have more page views on average, we only count the number of page views received up to 7 days after publication. Given that page views tend to sharply drop off soon after publication since recency is quite important to the News, the number of page views obtained during the 7 days after an article is published represents the vast majority (usually well over 90\%) of total page views an article receives. In total, our data consists of 248,161,455 page views. As seen in Figure 1 below, the distribution of page views is highly skewed with very heavy tails. After applying a log transformation (as seen in Figure 2), our distribution looks considerably more normal.

@@

In addition to aggregating the counts, when parser the internal web traffic data, we made sure to store various relevant article meta-data features such as the headline, time of publishing, authors, section, desk, and the NYT's internal article content description tags (if available\footnote{by this we mean that the NYT only produces these content tags only for some articles, its not a question of whether the data is available to us}).

\subsection{Parsed NYT Article Content Data}
Unfortunately, the NYT internal web traffic data does not the actual content displayed on each webpage, which is a very important aspect of our project. To resolve this issue, we used the python library ``newspaper". This library helped to scrape the html from the NYT website and extract all the raw textual data from the html. We further used some regular expressions to clean up what the library missed. In total, our 6682 articles contain 4,685,021 words of text.  We find that the distribution of article length, much like the distribution of page views, is highly skewed. We again apply a log transformation. Histograms and various summary statistics of for the distribution of article length and log article length are provided below:

@@

In addition to extracting the raw text data, we made sure to check if the articles contained non-textual content such as pictures or videos. We created indicator variables that denote the presence of such content within an article.

@@

In preparing to perform our regression, we construct a number of features, using both the parsed NYT article data and secondary sources. These features include the Flesch reading ease, the (guessed) gender of the author(s), the popularity of the author(s), the sentiment of the article text, the perplexity of the article text, and variables indicating the section the article appeared in and the article's content type. Found below is a full list of these features, as well as the methodology used to construct them and validation of the features, where appropriate.

\begin{description}
	\item[Flesch Reading Ease] One can conceive of a few hypotheses that relate the readership of content to the ease with which people can read it. Maybe more complicated pieces of text are more engaging, and are more likely to be read. On the other hand, perhaps pieces of text that are easier to read will be consumed by more people. In order to capture relationships such as these in our data, we include the Flesch reading ease. The Flesch reading ease is a metric developed by Flesch in 1948 \cite{flesch1948new}. The score indicates how difficult a piece of English text is to understand. Lower scores correspond to more difficult passages, and the highest score attainable is 120.0. The formula for calculating a passage's Flesch reading ease is

@@

	In order to measure something that will serve as a decent proxy for popularity, we programmatically searched for every distinct author in our dataset using Bing, and recorded the number of search results that were returned by the query. In cases where a particular article has more than one distinct author, we calculate an ``effective" popularity by simply calculating the average number of search results over all article authors.

	\item[Author Gender] We certainly hope that an author's gender would not have a causal impact on whether or not a particular article gets readership. However, we want to allow for this fact and measure this variable. We do so by constructing a feature that indicates the most likely gender of the article author(s). In cases where the gender of the author is unclear (e.g., Robin) or there are likely multiple authors with different genders (e.g., The New York Times Staff), we record a third gender value, ``ambiguous / unknown."

	We construct our gender data by cross-referencing the first names of all of the authors in our dataset against U.S. Social Security Administration baby name data from 1935 to 1997. If over 90\% of the babies with a given name have been male, we assume a given author is male. If over 90\% of the babies with a given name have been female, we assume a given author is female. Otherwise, we record ``ambiguous / unknown."

	\item[Material Type, Section, Desk, and Article Type] We also include dummy variables including the material type (e.g., `News' or `Obituary'), desk (e.g., `Weekend' or `Real Estate'), article type (`Blog post' or `Article'), and section (e.g., `Movies' or `World'). The hypothesis driving the decision to include these variables is that certain types of content (e.g., political news or international affairs) may be more widely read than local material (such as real estate) or less popular sections of the NY Times (e.g., the sports section).

@@

We also include features that attempt to capture the article sentiment and the article text perplexity. Since the construction of these features was more involved and involved validation of our algorithms, we discuss these two features in separate subsections.

@@

\noindent where $p_{ki}$ is the probability of a document of class $K$ including word $i$. Given that we transform this problem into log space, we can compute $\log(p(\mathbf{x}|C_k))$ as

@@

We coded up a basic implementation of the Naive Bayes algorithm, drawing heavy inspiration from Greg Lamp's 2014 python tutorial on Naive Bayes \cite{NaiveBayes}. In order to get the probabilities $p(C_k)$ and $p_{ki}$, we developed a training set of NY times data by providing 200 articles to workers on Amazon Mechanical Turk. Each turker was asked to score the sentiment toward the subject of the article in question on a scale from -2 to +2, with -2 being extremely negative and +2 being extremely positive. We recorded 5 scores for every article, and calculated the average sentiment reported by the Turkers. We classified any article having an average score greater than 0.5 as `positive'. Any article with an average sentiment less than -0.5 was `negative.' Any other articles were classified as `neutral.' Ultimately, our labels were 66\% neutral, 14.5\% negative, and 19.5\% positive. This is unsurprising, as a newspaper such as the New York Times likely strives for neutrality when reporting on most topics.

>>>>

\Authors{Dave Holtz\textsuperscript{1}, Jeremy Yang\textsuperscript{2}, Michael Zhao\textsuperscript{3*}} % Authors
\affiliation{\textsuperscript{1}\textit{dholtz@mit.edu}}
\affiliation{\textsuperscript{2}\textit{zheny@mit.edu}} 
\affiliation{\textsuperscript{3}\textit{mfzhao@mit.edu}}
\affiliation{\textsuperscript{*}\textit{Michael is using this project to fulfill Final Project requirements for both 6.864 as well as 6.867. Dave and Jeremy are only using this project for 6.867.}}

@@

\Abstract{Why do some webpages receive massive numbers of page views? To determine how content drives viewership, we construct a unique dataset of all articles published by the New York Times (NYT) in August 2013. Our dataset is built from 2 major components, the NYT's internal web traffic data and web content data extracted from the NYT website. We use the internal web traffic data to accurately track the number of page views of each article as well as construct a set of robust control variables such as the desk and section of each article. To build textual content features, we use various machine learning and statistical natural language processing techniques on our extracted content data to produce perplexity scores and to determine the sentiment and reading difficulty of each article. We also generate indicators that denote the presence of pictures, videos, etc. Lastly, we use some secondary data sources to build some additional control features such as estimated author gender and author popularity. We take all these features and build a predictive regression model. Overall, we do find that our textual features improve our predictive power, though rather modestly. However, due to time constraints, there remain many textual features we haven't yet implemented. We believe that adding these additional features may offer more significant improvements in performance.}

@@

In today's digital economy, both individuals and companies are very interested in attracting users to visit their websites in order to earn ad revenue. While many factors might motivate a user to visit a particular page, certainly one important factor is the content contained within that webpage. This paper explores the relationship between the content of a webpage and the number of page views it receives by constructing a unique dataset of all articles published by the New York Times (NYT) during August 2013. This dataset is built from two major components: the NYT's internal web traffic data and webpage content data parsed from the NYT website.

Typically, a study such as ours tends to be very difficult to conduct  accurate measures of viewership are either private of unavailable\footnote{While precise viewership data tends to be not available openly, oftentimes researchers use related observables, such as Facebook likes, or number of Tweets and Retweets}. Even in cases where page views are publically, for example Youtube or various other video sharing sights, feature extraction of the content is far too challenging given the tools we have available to us today\footnote{Though this is quickly changing with advances in computing power and development of better and better machine learning methods}. Fortunately, our access to the the NYT's internal web traffic data allows us to exactly measure the number of page views an article receives. The web traffic data is rather rich and also includes internal meta-data that we use to build various control features. Moreover, since we are working with mostly textual data, we are able to take advantage of recent advancements in machine learning and statistical NLP to do feature extraction on article text.

A similar study by Berger and Milkman (2012) \cite{berger2012makes} examines the relationship between content and word-of-mouth virality. They find that the emotional content of a NYT article is predictive of its virality. Using simple measures of an article's sentiment and emotionality, Berger and Milkman show that positive articles are more likely to show up on the New York Times "Most-Emailed" list. They also show that articles that evoke high physiological positive or negative arousal (such as awe or anger) tend to be more viral than articles that evoke deactivating emotions (sadness). We build on this study in two ways: first, we relate an article's content back to the number of page views it receives rather than its virality\footnote{Which companies arguably care more about since word-of-mouth virality is usually a means to increase page views}. Second, we employ more sophisticated machine learning feature extraction techniques.
%------------------------------------------------

\section{Data}
\subsection{NYT Internal Web Traffic Data}
Our NYT internal web traffic dataset is a record of all individual user activity on the NYT website covering the period of April 3rd, 2013 to October 31st, 2013. Each time a user\footnote{in this case, a ``user" is uniquely defined by a device/browser id. So while the same person might have multiple devices or may use multiple browsers, the NYT backend treats each device/browser combination as a unique ``user" even though in reality its all the same person. In some cases, we are able to link various id's together if the person happens to register an official user account on the NYT website and then logs into her account from multiple devices/browsers.} moves from one page to another on the NYT website, this activity is captured as an indivudal line json. This data is incredibly detailed and can potentially tell us who accessed what page when and from what location. In addition it tracks the duration a user stays on a webpage as well as the path a user took to get to her current webpage.\footnote{Big brother is watching you and we are big brother. Welcome to the age of Big Data} Overall, this data is over 20 terabytes in size and contains over 3 billion page views. It's important to note that not all page views are page views of content we care about. For example, some events that are also tracked in the data are searches or changes in user account settings. Since the scope of this dataset is so large, we initially restrict this project to a single month, August 2013, with plans of extending our analysis to our entire dataset for future work. We further restrict our dataset to only include articles or blogposts since these are the pieces of content that have contain readily extractable text data. For the sake of brevity, we use the term ``articles" to refer to both articles and blogposts unless we explicitly state otherwise.

We make a basic first pass through the data to simply to obtain a list of urls. After cleaning up the url data to ensure each url mapped uniquely to a particular piece of content, we were left with a total of 6682 urls. We then parse all the web data for the month of August and the first week of September count the number of impressions a url receives. In order to make an apples to apples comparison between articles we only count the number of page views received up to 7 days after publication since an article that's been out longer should have more page views in expectation. Given the tendency for the viewership of an article to drop off sharply soon after publication (as recency is important factor in determining of news readership), our 7-day measure generally represents the vast majority (well over 90\%) of total page views that an article receives\footnote{at least for a reasonable stretch of time}. Even after all this subsampling, our data still consists of 248,161,455 page views\footnote{Though one video by PSY completely crushes this number}. As seen in Figure 1 below, the distribution of page views is highly skewed with very heavy tails. After applying a log transformation (as seen in Figure 2), our distribution looks considerably more normal.

@@

In addition to aggregating the counts, when we parse the internal web traffic data, we make sure to extract various relevant article meta-data features such as the headline, time of publishing, authors, section, desk, and the NYT's internal article content description tags (if available).

\subsection{Parsed NYT Webpage Content Data}
Unfortunately, the NYT internal web traffic data does not contain the actual content displayed on each webpage, which is a very important aspect of our project. Luckily, all this content is freely hosted on the NYT website! Although the NYT limits the number of free articles you can access per month, the tracking system is cookie based. This means that scraping the raw html content via wget or something similar does not contribute to your article count limit\footnote{Alternatively, you can just keep clearing your cookies}. Specifically, we used the python library ``newspaper" to download the html content from the NYT website and then extract all the raw textual data from the html. We applied some additional regular expressions filters to clean up what the library missed. In total, our 6682 articles contain 4,685,021 words of text.  We find that the distribution of article length, much like the distribution of page views, is highly skewed and heavy tailed. We again apply a log transformation. Histograms and various summary statistics of for the distribution of article length and log article length are provided below:

@@

In addition to extracting the raw text data, we made sure to check if the articles also contained any addtional non-textual content such as pictures or videos. We created indicator variables that denote the presence of such content within an article.

@@

Using our collected this data, as well as some data from some additional secondary sources we construct the features that make up our predictive regression model. These features include the Flesch reading ease, the estimated gender of the author(s), the popularity of the author(s), variables indicating the section the article appeared in and the article's content type, the sentiment of the article text, and the perplexity of the article text. We provide a full list of these features below, as well as the methodology used to construct them. Where appropriate, we include discussion of testing and validation of our features and our algorithms.

\begin{description}
	\item[Flesch Reading Ease] One can conceive of a few competing hypotheses that relate the readership of content to the ease with which people can read it. Maybe more complicated pieces of text are more engaging, and are more likely to be read. On the other hand, perhaps pieces of text that are easier to read will be consumed by more people. In order to capture relationships such as these in our data, we include the Flesch reading ease. The Flesch reading ease is a metric developed by Flesch in 1948 \cite{flesch1948new}. The score indicates how difficult a piece of English text is to understand. Lower scores correspond to more difficult passages, and the highest score attainable is 120.0. The formula for calculating a passage's Flesch reading ease is

@@

	In order to measure something that will serve as a decent proxy for popularity, we programmatically searched for every distinct author in our dataset on Bing and recorded the number of search results that were returned by the query. In cases where a particular article has more than one distinct author, we calculate an ``effective" popularity by simply averaging number of search results over all article authors.

	\item[Author Gender] While we certainly don't think that an author's gender has a causal impact on the readership on an article. We believe that this feature allows us to control for some latent unobserved heterogeneity. For example, its not a stretch to think that the experiences of women authors are drastically different male authors and these differences are reflected in their writing. We construct this feature that indicates the most likely gender of the article author(s). In cases where the gender of the author is unclear (e.g., Robin) or there are likely multiple authors with different genders (e.g., The New York Times Staff), we record a third gender value, ``ambiguous / unknown."

	Our gender data is gathered by cross-referencing the first names of all of the authors in our dataset against U.S. Social Security Administration baby name data from 1935 to 1997. If over 90\% of the babies with a given name have been male, we assume a given author is male. If over 90\% of the babies with a given name have been female, we assume a given author is female. Otherwise, we record ``ambiguous / unknown."

	\item[Material Type, Section, Desk, and Article Type] We also include dummy variables including the material type (e.g., `News' or `Obituary'), publishing desk (e.g., `Weekend' or `Real Estate'), article type (`Blog post' or `Article'), and section (e.g., `Movies' or `World'). The hypothesis driving the decision to include these variables is that certain types of content (e.g., political news or international affairs) may be more widely read than local material (such as real estate) or less popular sections of the NY Times (e.g., the sports section).

@@

We also build features that attempt to capture the article sentiment and the article text perplexity. Since the construction of these features was considerably more complex and involved validation of our algorithms, we discuss these two features in separate subsections.

@@

\noindent where $p_{ki}$ is the probability a word $w_i$ conditional on a document of being of class $K$. Applying a log transformation, we can compute $\log(p(\mathbf{x}|C_k))$ as:

@@

We coded up a basic implementation of the Naive Bayes algorithm, drawing heavy inspiration from Greg Lamp's 2014 python tutorial on Naive Bayes \cite{NaiveBayes}. In order to get the probabilities $p(C_k)$ and $p_{ki}$, we needed some labeled training data. In order to obtain these labels, we selected a random subset of 200 articles from our dataset and created a task on Amazon Mechanical Turk. Each Turker was asked to score the sentiment toward the subject of the article in question on a scale from -2 to +2, with -2 being extremely negative and +2 being extremely positive. In order to make sure these scores were relatively robust, we recorded 5 scores for every article from 5 different Turkers and calculated the average sentiment score. We classified any article having an average score greater than 0.5 as `positive'. Any article with an average sentiment less than -0.5 was `negative.' Any other articles were classified as `neutral.' Ultimately, our labels were 66\% neutral, 14.5\% negative, and 19.5\% positive. This is unsurprising, as a newspaper such as the New York Times likely strives for neutrality when reporting on most topics.
